\chapter*{Acknowledgments}

First of all, I would like to thank Professor Marco Polverini who guided me along this difficult and long path. Before this research activity I had never studied Software Defined networks, except for a practical lesson held by him during the "Network Infrastructures" course. His help was essential to understand the complex concepts and technologies needed to complete this study. Various obstacles have arisen along the way and I am grateful that we have faced them together.\\

Honored to have known them during my studies, my thanks go to the extraordinary professors who have passed their knowledge to me with great passion. These have been years in which I feel I have learned a lot during the courses held by them. I wish anyone embarking on a course of study to be lucky enough to have such passionate and wise people as teachers.\\

I had the pleasure of sharing this journey with fantastic colleagues with whom I shared many beautiful moments. Their help and their company have been essential to overcome the obstacles of the course of study, it's always nice to have someone by your side.\\

To all my amazing friends who have accompanied me on this journey. The laughs and the funny moments we had together helped me and gave me strength in the worst moments. Thank you.\\

The most heartfelt thanks go to my family, and especially to my parents. 
"We are like dwarfs on the shoulders of giants" said the French philosopher Bernard of Chartres, and that is exactly how I feel. 
Theirs was not just help or support, they lifted me up on their shoulders and taught me to see the world with a different perspective.\\
Every day in which I felt weak and inadequate they reminded me to always look at the horizon, that even the sun of the most horrible day sets nevertheless and we will be given a new chance to shine again.\\
You have been amazing, and always will be.\\

\bigskip

Estote Semper Parati


\refstepcounter{chapter}